Extracting Function Signature from R Program Traces

### 2020-05-13 paper/main.tex
class discussion revisit

As discussed in Section~\ref{sec:R}, the R runtime has more than one notion
of type.  The {\it class} (rather, the {\it classes}, as values can have
multiple!) of function arguments are used by R to {\it dispatch} the correct
function call for the argument \AT{$\leftarrow$ wording}.  This is best
served with an example: \AT{Example, unless we talked about this earlier, in
  which case refer reader back.}

%
%
\subsubsection{Class Types}

The class of a value suggest a shape for the value, and we felt class to be a more succinct way of capturing struct-like types.

\AT{We used classes to stand-in for structs.  The discussion here is
  naturally tied with the discussion of structs, so we need to make sure to
  divide the discussion appropriately.}

>>>>

As discussed in Section~\ref{sec:R}, the R runtime has more than one notion of type.  
The classes of function arguments are used by R as part of its dispatch mechanism, the most prevalent of which (\AT{we have that number, too!}) is S3 dispatch, in which R dispatches on the (S3) class of the first argument to a function, selecting the appropriate method to call for the class.
To quickly illustrate, consider the code snippet in Figure~\ref{fig:dispatch-ex}.

\begin{lstlisting}
# Quick example of S3 dispatch.
\end{lstlisting}

\caption{S3 dispatch in R.}
\label{fig:dispatch-ex}
\end{center}
\end{figure}

Classes are very prevalent in R \AT{number}, and so we include a \code{class<class-name>} ... \AT{todo}
\typetracer was outfit to report the class of values, but some care needed to be taken so as to not overreport class types, as {\it all} values have a class in R.
The main classes we were interested in were matrix, array (higher dimensional matrices), and data.frames, in addition to any user-defined classes.
In many cases \AT{number?}, values with a user-defined class are in reality lists with named elements, or structs.
Including class types, and having \typetracer report them in place of structs, significantly reduced the noise in the analysis and allowed us to gather more focused data.
To get a sense for this difference, consider the code in Figure~\ref{fig:class-vs-struct}.

\begin{lstlisting}
# Example of gross struct that turns into class.
\end{lstlisting}

\caption{S3 dispatch in R.}
\label{fig:class-vs-struct}
\end{center}
\end{figure}

When set to ignore classes, \typetracer reports \AT{type} as the type of the function, as compared with \AT{other type} when configured to report classes.
